\thetitle

### 2020-12-04 sun_phases/spin_qudit_tomo.tex
removed redundant step from tomography protocol

As a special case, the phase-space representation $\rho^{\t{PS}}$ of a qudit state $\rho$ is common known as its Husimi distribution.
Performing tomography on an unknown qudit state $\rho$ is therefore equivalent to reconstructing the unknown distribution $\rho^{\t{PS}}$ on the sphere.
In principle, the representation $\rho^{\t{PS}}$ of a finite-dimensional qudit state $\rho$ can be reconstructed from the values $\rho^{\t{PS}}\p{\v v}=\bk{s_{\v v}|\rho|s_{\v v}}$ at a finite number of points $\v v$ on the sphere.
In practice, the value $\bk{s_{\v v}|\rho|s_{\v v}}$ is determined by measuring spin projection along $\v v$, which also provides measurement data on all spin projections $\bk{\mu_{\v v}|\rho|\mu_{\v v}}$ with $\mu\in\set{s,s-1,\cdots,-s}$ and $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$; one would like to make use of this additional data as well.
We clarify the connection between the quantum problem of reconstructing $\rho$ from spin projection measurements and the classical problem of reconstructing $\rho^{\t{PS}}$ from its values $\rho^{\t{PS}}\p{\v v}$ in the following section.
In addition to providing a new perspective on the problem of spin qudit tomography, the connection will allow us to place upper bounds on the minimal number of measurement axes $\v v$ necessary for full recovery of an unknown state.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
\begin{enumerate}
\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) by minimizing the quantum error scale $\epsilon_V$.
  If $\abs{V}$ is too large for such optimization to be practical, you can simply generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
\item For each axis $\v v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
\item Use the the estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ to compute the estimates
\begin{align}
  \tilde T_{\v v\ell,0}
  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\v v\mu}
\end{align}
of $\bk{T_{\v v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
\item For each $\ell\in\set{0,1,\cdots,d-1}$, build the measurement matrix $M_{V\ell}$ defined in Eq.~\eqref{eq:block}, and compute its singular value decomposition to identify the singular values $M^V_{\ell k}$ and corresponding right singular vectors $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$, indexed by an integer $k\in\set{1,2,\cdots,2\ell+1}$. % todo: cite codes
\item Compute the complex numbers
  \begin{align}
    \tilde\rho^V_{\ell k} \equiv \f1{M^V_{\ell k}} \sum_{\v v}
    x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0},
  \end{align}
  and the operators
  \begin{align}
    Q_{\ell k} \equiv \f1{M^V_{\ell k}} \sum_{\v v}
    x^V_{\ell k\v v} T_{\v v\ell,0},
  \end{align}
  and combine them into the estimate
  \begin{align}
    \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
    \approx \rho.
  \end{align}
\end{enumerate}
If reconstruction error is dominated by shot noise, then the total error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

>>>>

As a special case, the phase-space representation $\rho^{\t{PS}}$ of a qudit state $\rho$ is commonly known as its Husimi distribution.
Performing tomography on an unknown qudit state $\rho$ is therefore equivalent to reconstructing the unknown distribution $\rho^{\t{PS}}$ on the sphere.
In principle, the representation $\rho^{\t{PS}}$ of a finite-dimensional qudit state $\rho$ can be reconstructed from the values $\rho^{\t{PS}}\p{\v v}=\bk{s_{\v v}|\rho|s_{\v v}}$ at a finite number of points $\v v$ on the sphere.
In practice, the value $\bk{s_{\v v}|\rho|s_{\v v}}$ is determined by measuring spin projection along $\v v$, which also provides measurement data on all spin projections $\bk{\mu_{\v v}|\rho|\mu_{\v v}}$ with $\mu\in\set{s,s-1,\cdots,-s}$ and $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$; one would like to make use of this additional data as well.
We clarify the connection between the quantum problem of reconstructing $\rho$ from spin projection measurements and the classical problem of reconstructing $\rho^{\t{PS}}$ from its values $\rho^{\t{PS}}\p{\v v}$ in the following section.
In addition to providing a new perspective on the problem of spin qudit tomography, the connection will allow us to place upper bounds on the minimal number of measurement axes $\v v$ necessary for full recovery of an unknown state.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
\begin{enumerate}
\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) by minimizing the quantum error scale $\epsilon_V$ in Eq.~\eqref{eq:bound}.
  If $\abs{V}$ is too large for such optimization, you can simply generate many sets of random measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
  Note that computing the error scale $\epsilon_V$ requires, for each $\ell\in\set{0,1,\cdots,d-1}$, constructing the measurement matrix $M_{V\ell}$ in Eq.~\eqref{eq:block} and computing its singular value decomposition.
  Save all measurement matrix data associated with the final measurement axes $V$ for later use. % todo: cite codes
\item For each axis $\v v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
\item Use the the estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ to compute the estimates
\begin{align}
  \tilde T_{\v v\ell,0}
  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\v v\mu}
\end{align}
of $\bk{T_{\v v\ell,0}}_\rho$, where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding right singular vectors by $\v x^V_{\ell k} = \sum_{\v v} x^V_{\ell k \v v} \ket{\v v}$, compute the coefficients and operators
  \begin{align}
    \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
    x^V_{\ell k\v v} \, \tilde T_{\v v\ell,0}, \\
    Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\v v}
    x^V_{\ell k\v v} T_{\v v\ell,0},
  \end{align}
  and combine them into the estimate
  \begin{align}
    \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
    \approx \rho.
  \end{align}
\end{enumerate}
If reconstruction error is dominated by shot noise, then the total error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is approximately $\E_V\p{\tilde\rho_V}\approx\E_V\p{\rho}$, which can be computed from Eq.~\eqref{eq:error}. % todo: cite codes
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.
